Commutative Character Sheaves

### 2015-11-11 followup.tex
Started work on the proofs of the lemmata in Section 3.

@@

Now suppose $q : G \to Q$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
Q
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $Q$.
Consider the functor
\[
q^* : \CS(Q) \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(Q) \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(Q) \arrow{dl}{\text{comparison}}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

\begin{lemma}\label{lem:torsor}
If $q : G \to Q$ is an $H$-torsor in the fppf topology 
then the comparison functor $\CS(Q) \to \CS_H(G)$ is an equivalence.
\end{lemma}

@@

\begin{lemma}\label{lem:HH}
If $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ may be apprehended as a morphism of $H\times H$-equivariant linear character sheaves on $G\times G$; likewise $\vartheta : \cs{L}\boxtimes \cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$. 
\end{lemma}

@@

\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}]
The abelianization $G\ab\ceq G/G\der$ exists in schemes over $\Fq$ \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}.
Now,
\[
\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
1 \arrow{r} & G\der \arrow{r}{i} & G \arrow{r}{q} & G\ab \arrow{r} & 1
\end{tikzcd}
\]
gives
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{comparison}} \\
\arrow{u}{\text{forget}} \CS'(G)  \arrow{r}{\beta \mapsto \nu} &  \CS_{G\der}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]
%
Since the quotient $q : G \to G\ab$ is a trivial $G\der$-torsor in the fppf topology \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}, 
it now follows from Lemma~\ref{lem:torsor} that the comparison functor $\CS(G\ab)\to \CS_{G\der}(G)$ is an equivalence.
On the other hand, by Lemma~\ref{lem:beta}, $\CS'(G) \to \CS_{G\der}(G)$, defined by $(\cs{L},\mu,\beta)\mapsto (\cs{L},\mu, \nu)$, is an equivalence.
Since the essential image of the forgetful functor $\CS'(G)\to \CS(G)$ is precisely the category of $(\cs{L},\mu)\in \CS(G)$ such that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$, this completes the proof of Proposition~\ref{prop:Gder_triv}.
%that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ if and only if $\cs{L} = q^*\cs{L}\ab$ for some $\cs{L}\ab \in \CS(G\ab)$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{G\der}(G)$ is an equivalence, it determines a full subcategory $\CCS_{G\der}(G)$ which is equivalent to $\CCS(G\ab)$. 
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{} \\
\arrow{u}{\text{forget}} \CS'(G)  \arrow{r}{\beta \mapsto \nu} &  \CS_{G\der}(G) \arrow{u}{\text{forget}}  & \CCS(G\ab) \arrow{u} \arrow{dl} \\
\CCS(G) \arrow{u} \arrow{r}{\beta \mapsto \nu} & \arrow{u}  \CCS_{G\der}(G) & 
\end{tikzcd}
\]
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of the equivalence $\CS_{G\der}(G) \to \CS'(G)$ is $\CCS(G)$.
To that end, suppose $(\cs{L},\mu,\beta)\in \CS'(G)$ and $(\cs{L},\mu,\nu) = q^*(\cs{L}\ab,\mu\ab)$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{ssec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as in Section~\ref{ssec:noncom}.
Then the diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $(q\times q)^*$ to the diagram in Definition~\ref{def:CCScom}; 
in particular $\gamma = (q\times q)^* \xi$.

>>>>

\begin{lemma}\label{lem:HH}
If $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ is a morphism of $H\times H$-equivariant linear character sheaves on $G\times G$; likewise $\vartheta : \cs{L}\boxtimes \cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$. 
\end{lemma}

@@

Now suppose $q : G \to Q$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
Q
\end{tikzcd}
\]
Let $\cs{L}$ be a linear character sheaf on $Q$.
Consider the functor
\[
q^* : \CS(Q) \to \CS(G)
\]
given on objects by $(\cs{L},\mu) \mapsto (q^*\cs{L}, (q\times q)^* \mu)$; cf [Dictionary].
Then $q^*\cs{L}$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu : a^* q^* \cs{L} \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L} = (q\circ p)^*\cs{L} \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation defines the comparison functor
\[
q_H^* : \CS(Q) \to CS_H(G)
\]
and shows that the functor $q^* : \CS(Q) \to \CS(G)$ factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(Q) \arrow{dl}{q_H^*}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

\begin{lemma}\label{lem:torsor}
If $q : G \to Q$ is an $H$-torsor in the fppf topology 
then $q_H^*: \CS(Q) \to \CS_H(G)$ is an equivalence.
\end{lemma}

@@

@@

\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}]
The abelianization $G\ab\ceq G/G\der$ exists in schemes over $\Fq$ \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}.
Now,
\[
\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
1 \arrow{r} & G\der \arrow{r}{i} & G \arrow{r}{q} & G\ab \arrow{r} & 1
\end{tikzcd}
\]
gives
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{q_{G\der}^*} \\
\arrow{u}{\text{forget}} \CS'(G)  \arrow{r}{\beta \mapsto \nu} &  \CS_{G\der}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]
%
Since the quotient $q : G \to G\ab$ is a trivial $G\der$-torsor in the fppf topology \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}, 
it now follows from Lemma~\ref{lem:torsor} that the comparison functor $\CS(G\ab)\to \CS_{G\der}(G)$ is an equivalence.
On the other hand, by Lemma~\ref{lem:beta}, $\CS'(G) \to \CS_{G\der}(G)$, defined by $(\cs{L},\mu,\beta)\mapsto (\cs{L},\mu, \nu)$, is an equivalence.
Since the essential image of the forgetful functor $\CS'(G)\to \CS(G)$ is precisely the category of $(\cs{L},\mu)\in \CS(G)$ such that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$, this completes the proof of Proposition~\ref{prop:Gder_triv}.
%that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ if and only if $\cs{L} = q^*\cs{L}\ab$ for some $\cs{L}\ab \in \CS(G\ab)$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{G\der}(G)$ is an equivalence, it determines a full subcategory $\CCS_{G\der}(G)$ which is equivalent to $\CCS(G\ab)$. 
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{q_{G\der}^*} \\
\arrow{u}{\text{forget}} \CS'(G)  \arrow{r}{\beta \mapsto \nu} &  \CS_{G\der}(G) \arrow{u}{\text{forget}}  & \CCS(G\ab) \arrow{u} \arrow{dl} \\
\CCS(G) \arrow{u} \arrow{r}{\beta \mapsto \nu} & \arrow{u}  \CCS_{G\der}(G) & 
\end{tikzcd}
\]
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of the equivalence $\CS_{G\der}(G) \to \CS'(G)$ is $\CCS(G)$.
To that end, suppose $(\cs{L},\mu,\beta)\in \CS'(G)$ and $(\cs{L},\mu,\nu) = q^*(\cs{L}\ab,\mu\ab)$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{ssec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as in Section~\ref{ssec:noncom}.
Then the diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $(q\times q)^*$ to the diagram in Definition~\ref{def:CCScom}; 
in particular $\gamma = (q\times q)^* \xi$.
